A Roadmap for\\HEP Software and Computing R\&D\\for the 2020s

### 2018-11-12 CWP/papers/HSF-CWP-2017-01_roadmap/latex/cwp-main.tex
Small fixes to references and LaTeX formatting
Added OS X friendly like to .gitignore (.DS_Store is the HPFS metadata directory)

\begin{itemize}
\item
  Review the implementations of physics models, including the
  assumptions, approximations, and
  limitations. In the best cases this can achieve higher precision and
  improve runtime performance through code modernisation\cite{Novak2018}.
  The extension of the validity of models up to energies of 100 TeV is foreseen
  for the Future Circular Collider (FCC) project~\cite{FCC} and provides
  a good opportunity for this modernisation.
\item
  Redesigning, developing, and commissioning detector simulation
  toolkits to be more efficient when executed on current vector CPUs and
  emerging new architectures, including GPUs, where use of SIMD
  vectorisation is vital; this includes porting and optimising the
  experiments' simulation applications to allow exploitation of large
  HPC facilities.
\item
  Exploring different Fast Simulation options, where the full detector
  simulation is replaced, in whole or in part, by computationally
  efficient techniques. An area of investigation is common frameworks
  for fast tuning and validation.
\item
  Developing, improving and optimising geometry tools that can be shared
  am\-ong experiments to make the modeling of complex detectors
  computationally more efficient, modular, and transparent.
\item
  Developing techniques for background modeling, including contributions
  of multiple hard interactions overlapping the event of interest in
  collider experiments (pileup).
\item
  Revisiting digitisation algorithms to improve performance and
  exploring opportunities for code sharing among experiments.
\end{itemize}

@@

An important new development in the analysis domain has been the
emergence of new analysis tools coming from industry and open-source
projects (e.g. Jupyter notebooks~\cite{Jupyter}, the scikit-learn package~\cite{scikit-learn}), and this presents new opportunities for improving the HEP
analysis software ecosystem. The HEP community is very interested in
using these software tools, together with established components, in an
interchangeable way. The main challenge will be to enable new
open-source tools to be plugged in dynamically to the existing ecosystem
and to provide mechanisms that allow the existing and new components to
interact and exchange data efficiently. To improve our ability to
analyse much larger datasets, R\&D will be needed to investigate file
formats, compression algorithms, and new ways of storing and accessing
data for analysis and to adapt workflows to run on future computing
infrastructures.

@@

One new model of data analysis, developed outside of HEP, maintains the
concept of sequential reduction, but mixes interactivity with
batch processing. These exploit new cluster management systems, most

notably Apache Spark~\cite{Spark,Armbrust:2015:SSR:2723372.2742797}, which uses open-source tools contributed both by
industry and the data-science community. Other products implementing the
same analysis concepts and workflows are emerging, such as TensorFlow~\cite{tensorflow2015-whitepaper},
Dask~\cite{rocklin2015dask,dask_ref}, Pachyderm~\cite{thepachydermteam}, 
Blaze~\cite{wiebe2014blaze}, Parsl~\cite{babuji_yadu_2017_853492}, 
and Thrill~\cite{bingmann2016thrill}. This approach can complement
the present and widely adopted Grid processing of datasets. It may

@@

The use of ML in HEP analyses has become commonplace over the past two
decades, and the most common use case has been in signal/background
classification. The vast majority of HEP analyses published in recent
years have used the HEP-specific software package TMVA~\cite{TMVA} included in ROOT. Recently, however, many HEP analysts have begun
migrating to non-HEP ML packages such as scikit-learn~\cite{scikit-learn}
and Keras~\cite{Keras}, although these efforts have yet to result
in physics publications from major collaborations. Data scientists at
Yandex created a Python package that provides a consistent API to most
ML packages used in HEP~\cite{REP}. Packages like Spearmint~\cite{Spearmint} and scikit-optimize~\cite{scikit-optimize} perform Bayesian optimisation and can improve HEP Monte
Carlo work.

>>>>

\begin{itemize}
\item
  Review the implementations of physics models, including the
  assumptions, approximations, and
  limitations. In the best cases this can achieve higher precision and
  improve runtime performance through code modernisation~\cite{Novak2018}.
  The extension of the validity of models up to energies of 100 TeV is foreseen
  for the Future Circular Collider (FCC) project~\cite{FCC} and provides
  a good opportunity for this modernisation.
\item
  Redesigning, developing, and commissioning detector simulation
  toolkits to be more efficient when executed on current vector CPUs and
  emerging new architectures, including GPUs, where use of SIMD
  vectorisation is vital; this includes porting and optimising the
  experiments' simulation applications to allow exploitation of large
  HPC facilities.
\item
  Exploring different Fast Simulation options, where the full detector
  simulation is replaced, in whole or in part, by computationally
  efficient techniques. An area of investigation is common frameworks
  for fast tuning and validation.
\item
  Developing, improving and optimising geometry tools that can be shared
  am\-ong experiments to make the modeling of complex detectors
  computationally more efficient, modular, and transparent.
\item
  Developing techniques for background modeling, including contributions
  of multiple hard interactions overlapping the event of interest in
  collider experiments (pileup).
\item
  Revisiting digitisation algorithms to improve performance and
  exploring opportunities for code sharing among experiments.
\end{itemize}

@@

An important new development in the analysis domain has been the
emergence of new analysis tools coming from industry and open-source
projects (e.g. Jupyter notebooks~\cite{Jupyter}, the scikit-learn package~\cite{Pedregosa:2011:SML:1953048.2078195}), and this presents new opportunities for improving the HEP
analysis software ecosystem. The HEP community is very interested in
using these software tools, together with established components, in an
interchangeable way. The main challenge will be to enable new
open-source tools to be plugged in dynamically to the existing ecosystem
and to provide mechanisms that allow the existing and new components to
interact and exchange data efficiently. To improve our ability to
analyse much larger datasets, R\&D will be needed to investigate file
formats, compression algorithms, and new ways of storing and accessing
data for analysis and to adapt workflows to run on future computing
infrastructures.

@@

One new model of data analysis, developed outside of HEP, maintains the
concept of sequential reduction, but mixes interactivity with
batch processing. These exploit new cluster management systems, most
notably Apache Spark~\cite{Spark,Armbrust:2015:SSR:2723372.2742797}, which uses open-source tools contributed both by
industry and the data-science community. Other products implementing the
same analysis concepts and workflows are emerging, such as TensorFlow~\cite{tensorflow2015-whitepaper},
Dask~\cite{rocklin2015dask,dask_ref}, Pachyderm~\cite{thepachydermteam},
Blaze~\cite{wiebe2014blaze}, Parsl~\cite{babuji_yadu_2017_853492},
and Thrill~\cite{bingmann2016thrill}. This approach can complement
the present and widely adopted Grid processing of datasets. It may

@@

The use of ML in HEP analyses has become commonplace over the past two
decades, and the most common use case has been in signal/background
classification. The vast majority of HEP analyses published in recent
years have used the HEP-specific software package TMVA~\cite{TMVA} included in ROOT. Recently, however, many HEP analysts have begun
migrating to non-HEP ML packages such as scikit-learn~\cite{Pedregosa:2011:SML:1953048.2078195}
and Keras~\cite{Keras}, although these efforts have yet to result
in physics publications from major collaborations. Data scientists at
Yandex created a Python package that provides a consistent API to most
ML packages used in HEP~\cite{REP}. Packages like Spearmint~\cite{Spearmint} and scikit-optimize~\cite{scikit-optimize} perform Bayesian optimisation and can improve HEP Monte
Carlo work.
